---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# kpitools

`r badger::badge_custom("dev version", as.character(packageVersion("kpitools")), "blue", "https://github.com/CTU-Bern/kpitools")`
[![R-CMD-fullcheck](https://github.com/CTU-Bern/kpitools/actions/workflows/R-CMD-full.yaml/badge.svg)](https://github.com/CTU-Bern/kpitools/actions/workflows/R-CMD-full.yaml)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Tools for creating key performance indicator (KPI) reports.

# Example usage 

The package can be installed from [github](https://github.com/CTU-Bern/kpitools) via the `remotes` package

```{r gh-installation, eval = FALSE}
# install.packages("remotes")
remotes::install_github("CTU-Bern/kpitools")
```

Note that `remotes` treats any warnings (e.g. that a certain package was built under a different version of R) as errors. If you see such an error, run the following line and try again:

```{r remotes-error, eval = FALSE}
Sys.setenv(R_REMOTES_NO_ERRORS_FROM_WARNINGS = "true")
```

And loaded via
```{r, message=FALSE}
library(kpitools)
```

The main function is the `kpi` function. A dataframe is passed to it together with the `var`iable that is of interest for the current KPI. A summary function also needs to be passed which determines how the KPI is calculated.
```{r, eval = TRUE}
data(mtcars)

mtcars$highmpg <- mtcars$mpg > 20
```


```{r}
kpis <- (mtcars %>%
  kpi(var = "highmpg",                         # variable to be summarized (focus of the KPI)  
      kpi_fn = kpi_fn_perc,                    # summary function   
      txt = "Percentage MPG > 20",             # (optional) nicer text to add to tables 
      by = "cyl",                               # (optional) stratifying variable 
      cutpoints = c(0,33.3,66.6,100),          # (optional) cutoff points 
      cutlabels = c("Low", "Medium", "High"))) # (optional) labels for the cutoff points

```

There is a plot method for the output from `kpi` which returns a list of `ggplot2` objects.
```{r, fig.height=1.5}
plot <- plot(kpis)
plot$cyl +
  theme_kpitools()
```


For further details, see the vignette:
```{r, eval = FALSE}
vignette("kpitools")
```



