---
title: "Using `kpitools`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Using `kpitools`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`kpitools` is intended to aid in creating reports on key performance indicators 
(KPIs). KPIs must be defined by the user, which can be a substantial undertaking, 
but `kpitools` then takes away some of the pain of putting them into a report.

`kpitools` is loaded as any other package, with the `library` or require `functions`:


```{r setup, message=FALSE}
library(kpitools)
```


Once the KPIs are defined conceptually, it should be clear what type of indicator one needs to
report. For instance, using the `mtcars` dataset, we may have an indicator for 
mileage. We're not really interested in the individual cars, but at the group of 
products as a whole. If the mean mileage is lower than 15, we may have to take 
actions to reduce the proportion. In a clinical trial, such remedial actions might 
be to retrain staff or, in a particularly serious case, closing a site completely.

To use `kpitools`, we must first load the data and create the indicators themselves
(using whatever method for a given dataset). To begin with, we will start with 
the `mpg` variable in the `mtcars` data.

```{r}
data(mtcars)
```

`kpitools` produces summary information on the defined KPIs. We 
tell the `kpi` function which dataset to use, which variable to summarize and how
to summarize it (`kpi` summarizes using the mean by default).

```{r}
x <- (kpi(data = mtcars, 
          var = "mpg"))
x
```


We see that the `r x$overall$calc$N` cars have a mean MPG of `r round(x$overall$calc$stat)` miles per gallon.

We can improve the output slightly by adding a more meaningful label via the `txt` argument.


```{r}
kpi(data = mtcars, 
    var = "mpg",
    txt = "Miles per gallon")

```

## Summary functions

`kpitools` comes with a set of functions for summarizing the KPIs.  

|Function           |Description
|:------------------|:---------------------------------------
|*Categorical KPIs* |
|`kpi_fn_n`         |takes the sum of objects
|`kpi_fn_prop`      |calculates the proportion of 1s
|`kpi_fn_perc`      |as `kpi_fn_prop` but on percentage scale
|*Numeric KPIs*     |
|`kpi_fn_median`    |calculates the median
|`kpi_fn_mean`      |calculates the mean
|`kpi_fn_iqr`       |calculates the interquartile range
|`kpi_fn_min`       |calculates the minimum
|`kpi_fn_max`       |calculates the maximum

These functions are passed to `kpi` via the `kpi_fn` argument. 

Say our KPI is on the proportion of cars with an MPG less than 15... first we create the variable and pass that to `kpi` with the appropriate summary function. 

```{r}
mtcars$mpg15 <- mtcars$mpg < 15
kpi(data = mtcars, 
    var = "mpg15",
    kpi_fn = kpi_fn_perc,
    txt = "Milage < 15 MPG")

```

### Using custom functions

Custom functions can also be utilized. Simply define a function which accepts a 
dataframe and performs the summary function of choice on a `var` variable. The summarised 
variable should be called `stat`. below is example to summarise the variable by
the variance.

```{r}
fn <- function(x){
  x %>%
    summarize(stat = var(var))
}

kpi(mtcars, "mpg", kpi_fn = fn)

```

We recommend adding an `N` the variable with the number of observations, which 
is used for plotting (the size of points is relative to the number of observations 
in the group).

```{r}
fn <- function(x){
  x %>%
    summarize(N = n(),
              stat = var(var))
}

kpi(mtcars, "mpg", kpi_fn = fn)

```


## Hierarchies

We can also look at specific subgroups (e.g. in a clinical trial, we would be 
interested in the performance of particular countries or centers) via the `by` 
argument. Using the `mtcars` example, we can use the number of cylinders.

```{r}
x_by <- (kpi(data = mtcars, 
             var = "mpg15",
             kpi_fn = kpi_fn_perc,
             txt = "Milage < 15 MPG",
             by = "cyl"))
x_by
```

Here we can see that all of the low mileage cars are in the 8 cylinder group, 
with `r x_by$cyl$calc$stat[3]`% of cars having low mileage.

We can also pass multiple variables to `by` (so that we can consider two levels 
or more levels of hierarchy simultaneously)
```{r}
x_by <- (kpi(data = mtcars, 
             var = "mpg15",
             kpi_fn = kpi_fn_perc,
             txt = "Milage < 15 MPG",
             by = c("cyl", "am")))
x_by
```

Here, we see that cars with manual gearboxes are more problematic than those with
automatic gearboxes.

## Plotting
Where hierarchies have been used, KPIs can also be visualized by plotting the 
object, which returns a list of 
`ggplot2` objects (so that you can modify the plots afterwards; we suggest a wide, 
short plot, e.g. `fig.width=7, fig.height=1.5` in the R chunk header). 

```{r, fig.width=7, fig.height=1.5}
(p <- plot(x_by))
```

The size of the points is relative to the number of observations (participants)
in the site/group.

`kpitools` contains a `ggplot2` which can be used to remove the y-axis, move the 
legend and make various other changes:
```{r, fig.width=7, fig.height=1.5}
p[[1]] + theme_kpitools()
```

Modifications can be made to all plots easily via e.g. `lapply` or `purrr::map`

```{r}
p <- purrr::map2(p, names(p), function(x, y){
  x +                                       # extract the ggplot object
    theme_kpitools() +                      # apply the theme
    guides(size = guide_legend(nrow = 1)) + # make other modifications
    ggtitle(toupper(y))                     # add a title
})
```

The `patchwork` package, for example, could then be used to arrange the plots 

```{r, fig.width=7, fig.height=3}
patchwork::wrap_plots(p, ncol = 1)
```


## From statistics to risk

It is quite common when talking about KPIs to link a given value of the summary 
statistic to a degree of risk. `kpitools` provides provisions for this too, via 
the `breakpoints` argument.

```{r}
kpi(data = mtcars, 
    var = "mpg",
    txt = "Milage per gallon",
    by = "cyl",
    breakpoints = c(0, 20, 30))
```




